---
title: "SimulationStudySummary"
author: "Xiang Ji"
date: "9/7/2017"
output:
html_document: default
pdf_document: default
---

This R markdown file summarizes Simulation Study results. 

```{r, echo=FALSE}
rm(list=ls())  # clean up workspace
setwd("/Users/xji3/GitFolders/YeastIGCTract/SimulationStudy/")

Tract.list <- c(3.0, 10.0, 50.0, 100.0, 200.0, 300.0, 400.0, 500.0)
# First read in HMM results
# from summary file
for(tract in Tract.list){
  hmm.tract.summary <- NULL
  for(sim in 1:100){
    hmm.summary <- paste("./summary/Tract_", toString(tract), '.0/sim_', 
                         toString(sim), '/HMM_YDR418W_YEL054C_MG94_nonclock_sim_',
                         toString(sim), '_1D_summary.txt', sep = "")
    if (file.exists(hmm.summary)){
      all <- readLines(hmm.summary, n = -1)
      col.names <- paste("sim_", toString(sim), sep = "")
      row.names <- strsplit(all[length(all)], ' ')[[1]][-1]
      summary_mat <- as.matrix(read.table(hmm.summary, 
                                          row.names = row.names, 
                                          col.names = col.names))
      hmm.tract.summary <- cbind(hmm.tract.summary, summary_mat)     
    }
    
  }
  assign(paste("HMM_Tract_", toString(tract), "_summary", sep = ""), hmm.tract.summary)
}

# from plots
for(tract in Tract.list){
  hmm.tract.plots <- NULL
  for(sim in 1:100){
    hmm.plot <- paste("./plot/Tract_", toString(tract), '.0/sim_', 
                      toString(sim), '/HMM_YDR418W_YEL054C_lnL_sim_',
                      toString(sim), '_1D_surface.txt', sep = "")
    if (file.exists(hmm.plot)){
      lnL.surface <- read.table(hmm.plot)
      max.idx <- which.max(lnL.surface[, 2])
      new.summary <- matrix(c(3.0*exp(-lnL.surface[max.idx, 1]), lnL.surface[max.idx, 2]), 2, 1)
      rownames(new.summary) <- c("tract in nt", "lnL")
      colnames(new.summary) <- paste("sim_", toString(sim), sep = "")
      hmm.tract.plots <- cbind(hmm.tract.plots, new.summary)     
    }
  }
  assign(paste("HMM_Tract_", toString(tract), "_plot", sep = ""), hmm.tract.plots)
}

# Now read in actual mean tract length in each simulated dataset
for (tract in Tract.list){
  sim.tract <- NULL
  for(sim in 1:100){
    sim_log <- paste("./Tract_", toString(tract), ".0_HKY/sim_", toString(sim), 
                     "/YDR418W_YEL054C_sim_", toString(sim), "_IGC.log", sep = "")
    # now read in log file
    log_info <- read.table(sim_log, header = TRUE)
    realized.tract.length <- log_info[, "stop_pos"] - log_info[, "start_pos"] + 1
    potential.tract.length <- log_info[, "tract_length"]
    # Now get longest length of variant subtract of each tract
    diff.tracts <- log_info[, "num_diff"] > 0
    if (sum(diff.tracts)){
      subtract.length.list <- NULL
      for(row.num in (1:dim(log_info)[1])[diff.tracts]){
        donor.seq <- strsplit(toString(log_info[row.num, "template_seq"]), "")[[1]]
        recipient.seq <- strsplit(toString(log_info[row.num, "overide_seq"]), "")[[1]]
        first.pos <- FALSE
        for(seq.pos in 1:length(donor.seq)){
          if(donor.seq[seq.pos] != recipient.seq[seq.pos]){
            last.pos <- seq.pos
            if(!first.pos){
              first.pos <- seq.pos
            }
          }
        }
        subtract.length.list <- c(subtract.length.list, last.pos - first.pos + 1)
      }
    }
    new.info <- matrix(c(dim(log_info)[1], mean(potential.tract.length), sd(potential.tract.length), 
                         mean(realized.tract.length), sd(realized.tract.length),
                         mean(potential.tract.length[diff.tracts]), 
                         mean(realized.tract.length[diff.tracts]),
                         mean(subtract.length.list),
                         sum(log_info[, "num_diff"] > 1)),
                       9, 1)
    rownames(new.info) <- c("num IGC", "mean potential tract length", "sd potential tract length", 
                            "mean realized tract length", "sd realized tract length", 
                            "mean potential nonidentical tract length", "mean realized nonidentical tract length",
                            "mean subtract length", "num IGC with at least two variant sites")
    colnames(new.info) <- paste("sim_", toString(sim), sep = "")
    sim.tract <- cbind(sim.tract, new.info)
  }
  assign(paste("sim.tract.", toString(tract), sep = ""), sim.tract)
}

guess.list <- c(50.0, 100.0, 250.0, 500.0)
# Now read in PSJS summary results 
for(tract in Tract.list){
  for(guess in guess.list){
    PSJS.tract.summary <- NULL
    for(sim in 1:100){
      
      PSJS.summary <- paste("./summary/Tract_", toString(tract), '.0_HKY/sim_', 
                            toString(sim), '/PSJS_HKY_rv_sim_',
                            toString(sim), "_Tract_", toString(tract), '.0_guess_',
                            toString(guess),'.0_nt_summary.txt', sep = "")
      if (file.exists(PSJS.summary)){
        all <- readLines(PSJS.summary, n = -1)
        col.names <- paste("sim_", toString(sim), sep = "")
        row.names <- strsplit(all[length(all)], ' ')[[1]][-1]
        summary_mat <- as.matrix(read.table(PSJS.summary, 
                                            row.names = row.names, 
                                            col.names = col.names))
        PSJS.tract.summary <- cbind(PSJS.tract.summary, summary_mat)
      }
    }
    assign(paste("PSJS_HKY_Tract_", toString(tract), "_guess_", 
                 toString(guess), "_summary", sep = ""), PSJS.tract.summary)
  }
}

# Now combine all initial guess results
for(tract in Tract.list){
  combined.PSJS.tract.summary <- NULL
  col.list <- NULL
  for ( sim_num in 1:100){
    sim_col <- paste("sim_", toString(sim_num), sep = "")
    best.lnL <- -Inf
    best.guess <- NULL
    for(guess in guess.list){
      target_summary <- get(paste("PSJS_HKY_Tract_", toString(tract), "_guess_", toString(guess), "_summary", sep = ""))
      if(sim_col %in% colnames(target_summary) ){
        if (target_summary["ll", sim_col] > best.lnL){
          best.lnL <- target_summary["ll", sim_col]
          best.guess <- guess        
        }
      }
    }
    if(! is.null(best.guess)){
      combined.PSJS.tract.summary <- cbind(combined.PSJS.tract.summary, 
                                           get(paste("PSJS_HKY_Tract_", toString(tract), "_guess_", toString(best.guess), "_summary", sep = ""))[, sim_col]) 
      col.list <- c(col.list, sim_col)
    }
    
  }
  colnames(combined.PSJS.tract.summary) <- col.list
  assign(paste("PSJS_HKY_Tract_", toString(tract), "_combined_summary", sep = ""), combined.PSJS.tract.summary)
}

# Now read in PSJS summary results of Old Simulated Datasets 
for(tract in Tract.list){
  for(guess in guess.list){
    PSJS.tract.summary <- NULL
    for(sim in 1:100){
      
      PSJS.summary <- paste("./summary/IGCgeo_", toString(tract), '.0/sim_', 
                            toString(sim), '/PSJS_HKY_rv_sim_',
                            toString(sim), "_IGCgeo_", toString(tract), '.0_guess_',
                            toString(guess),'.0_summary.txt', sep = "")
      if (file.exists(PSJS.summary)){
        all <- readLines(PSJS.summary, n = -1)
        col.names <- paste("sim_", toString(sim), sep = "")
        row.names <- strsplit(all[length(all)], ' ')[[1]][-1]
        summary_mat <- as.matrix(read.table(PSJS.summary, 
                                            row.names = row.names, 
                                            col.names = col.names))
        PSJS.tract.summary <- cbind(PSJS.tract.summary, summary_mat)
      }
    }
    assign(paste("Old_PSJS_HKY_Tract_", toString(tract), "_guess_", 
                 toString(guess), "_summary", sep = ""), PSJS.tract.summary)
  }
}

# Now combine all initial guess results using old simulated data
for(tract in Tract.list){
  combined.PSJS.tract.summary <- NULL
  col.list <- NULL
  for ( sim_num in 1:100){
    sim_col <- paste("sim_", toString(sim_num), sep = "")
    best.lnL <- -Inf
    best.guess <- NULL
    for(guess in guess.list){
      target_summary <- get(paste("Old_PSJS_HKY_Tract_", toString(tract), "_guess_", toString(guess), "_summary", sep = ""))
      if(sim_col %in% colnames(target_summary) ){
        if (target_summary["ll", sim_col] > best.lnL){
          best.lnL <- target_summary["ll", sim_col]
          best.guess <- guess        
        }
      }
    }
    if(! is.null(best.guess)){
      combined.PSJS.tract.summary <- cbind(combined.PSJS.tract.summary, 
                                           get(paste("Old_PSJS_HKY_Tract_", toString(tract), "_guess_", toString(best.guess), "_summary", sep = ""))[, sim_col]) 
      col.list <- c(col.list, sim_col)
    }
    
  }
  colnames(combined.PSJS.tract.summary) <- col.list
  assign(paste("Old_PSJS_HKY_Tract_", toString(tract), "_combined_summary", sep = ""), combined.PSJS.tract.summary)
}
```

OK, Now show the performance summary

```{r, echo=FALSE}
# # HMM results
# for (tract in Tract.list){
#   # show how many stuck at boundary 1000 nt first
#   print(paste("Tract = ", toString(tract), sep = ""))
#   summary_mat <- get(paste("HMM_Tract_", toString(tract), "_plot", sep = ""))
#   
#   # histogram of inferred tract length
#   hist(summary_mat[1, summary_mat[1, ] < 999.], main = paste("Tract = ", toString(tract), sep = ""))
#   print(paste("Among total 100 simulated data sets, ", toString(sum(summary_mat[1, ] > 999)), 
#               " datasets stuck at 1000", sep = ""))
#   print(matrix(c("mean", mean(summary_mat[1, summary_mat[1, ] < 999.]), 
#                  "sd", sd(summary_mat[1, summary_mat[1, ] < 999.])), 2, 2))
# }
# 
# # PSJS results
# for(tract in Tract.list){
#   summary_mat <- get(paste("PSJS_HKY_Tract_", toString(tract), "_combined_summary", sep = ""))
#   # histogram of inferred tract length
#   hist(summary_mat["tract_length", ], main = paste("Tract = ", toString(tract), sep = ""))
#   print(matrix(c("mean", mean(summary_mat["tract_length", ]), 
#                  "sd", sd(summary_mat["tract_length", ])), 2, 2))  
# }
# 
# # exclude suspiciously long inferred tract length, plot again
# for(tract in Tract.list){
#   summary_mat <- get(paste("PSJS_HKY_Tract_", toString(tract), "_combined_summary", sep = ""))
#   # histogram of inferred tract length
#   hist(summary_mat["tract_length", summary_mat["tract_length", ] < 1000.0], main = paste("Tract = ", toString(tract), sep = ""))
#   print(matrix(c("mean", mean(summary_mat["tract_length", summary_mat["tract_length", ] < 1000.0]), 
#                  "sd", sd(summary_mat["tract_length", summary_mat["tract_length", ] < 1000.0])), 2, 2))  
# }
```

A plot of HMM surface that infer tract length at boundary from each tract length condition

```{r, echo=FALSE}
# # plot the first two datasets that HMM inferred tract length stuck at boundary of 1000.0 
# for (tract in Tract.list){
#   print(paste("Tract = ", toString(tract), sep = ""))
#   summary_mat <- get(paste("HMM_Tract_", toString(tract), "_plot", sep = ""))
#   count = 0
#   for(iter in 1:99){
#     if(summary_mat[1, iter] > 999. & count < 2){
#       to.plot <- read.table(paste("./plot/Tract_", toString(tract), ".0/", colnames(summary_mat)[iter], 
#                                   "/HMM_YDR418W_YEL054C_lnL_", colnames(summary_mat)[iter],
#                                   "_1D_surface.txt", sep = ""))
#       plot(3.0*exp(-to.plot[, 1]), to.plot[, 2], main = paste("Tract = ", toString(tract), " ",colnames(summary_mat)[iter], sep = "" ), type = "l", xlab = "tract length in nucleotides", 
#            ylab = "lnL")
#       count = count + 1
#     }
#   }
# }

```

Now plot the two plots of lnL
```{r, echo=FALSE}
# library(ggplot2)
# # Tract length = 10
# hmm.plot <- read.table("./plot/Tract_10.0/sim_1/HMM_YDR418W_YEL054C_lnL_sim_1_1D_surface.txt")
# PSJS.plot <- read.table("./plot/Tract_10.0/sim_1/PSJS_HKY_rv_sim_1_Tract_10.0_lnL_1D_surface.txt")
# ggplot(mapping = aes(x = 3.0*exp(-hmm.plot[,1]), y = hmm.plot[, 2], colour = "blue")) + geom_line() +
#   geom_line(aes(x = exp(-PSJS.plot[,1]), y = PSJS.plot[, 2]/488 + min(hmm.plot[, 2]) - min(PSJS.plot[, 2]/488),
#                 colour = "red")) +
#   xlab("Tract length in nucleotides") + 
#   ylab("lnL") + 
#   ggtitle("Tract length = 10, simulated dataet 1")
# 
# hmm.plot <- read.table("./plot/Tract_10.0/sim_100/HMM_YDR418W_YEL054C_lnL_sim_100_1D_surface.txt")
# PSJS.plot <- read.table("./plot/Tract_10.0/sim_100/PSJS_HKY_rv_sim_100_Tract_10.0_lnL_1D_surface.txt")
# ggplot(mapping = aes(x = 3.0*exp(-hmm.plot[,1]), y = hmm.plot[, 2], colour = "blue")) + geom_line() +
#   geom_line(aes(x = exp(-PSJS.plot[,1]), y = PSJS.plot[, 2]/488 + min(hmm.plot[, 2]) - min(PSJS.plot[, 2]/488),
#                 colour = "red")) +
#   xlab("Tract length in nucleotides") + 
#   ylab("lnL") + 
#   ggtitle("Tract length = 10, simulated dataet 100")
```

Now see how estimates from the two approaches differ from the actual mean tract length in each simulated data set.

```{r, echo=FALSE}
# for(tract in Tract.list){
#   sim.info <- get(paste("sim.tract.", toString(tract), sep = ""))
#   # Show mean and sd
#   print(matrix(c("potential mean", mean(sim.info["mean potential tract length", ]),
#           "realized mean", mean(sim.info["mean realized tract length", ]),
#           "geometric mean", tract,
#           "potential sd", mean(sim.info["sd potential tract length",], na.rm = TRUE),
#           "realized sd", mean(sim.info["sd realized tract length",], na.rm = TRUE),
#           "geometric sd", sqrt(tract^2-tract*3.0)), 2, 6))
#   hmm.info <- get(paste("HMM_Tract_", toString(tract), "_plot", sep = ""))
#   PSJS.info <- get(paste("PSJS_Tract_", toString(tract), "_summary", sep = ""))
#   shared.col <- intersect(colnames(hmm.info), colnames(PSJS.info))
#   
#   # Now show the ratio of HMM estimated tract / actual mean tracts in simulation
#   hmm.ratio <- hmm.info[1, shared.col]/sim.info[1, shared.col]
#   hist(hmm.ratio, main = paste("HMM ratio Tract = ", toString(tract), sep = ""))
#   print(matrix(c("HMM mean", mean(hmm.ratio), "HMM sd", sd(hmm.ratio)), 2, 2))
#   
#   # Now show the ratio of PSJS estimated tract / actual mean tracts in simulation
#   PSJS.ratio <- PSJS.info["tract_length", shared.col]/sim.info[1, shared.col]
#   hist(PSJS.ratio, main = paste("PSJS ratio Tract = ", toString(tract), sep = ""))
#   print(matrix(c("PSJS mean", mean(PSJS.ratio), "PSJS sd", sd(PSJS.ratio)), 2, 2))
#   
# }


```

####09152017

Show the PSJS estimated results for Simulated datasets with estimated tau

```{r}
realized.tract.dist <- function(L, p){
  x <- 1:(L-1)
  dist.1 <- (2 + (L-1-x)*p)/(1.0/p + L -1.)*(1-p)^(x-1)
  dist.L <- 1/p/(1.0/p + L -1.)*(1-p)^(L-1)
  dist <- c(dist.1, dist.L)
  mean.L <- sum(x * dist.1) + L*dist.L
  var.L <- sum(x^2 * dist.1) + L^2*dist.L - mean.L^2
  return(list(dist = dist, mean = mean.L, sd = sqrt(var.L)))
}

```

```{r}
Tract.list <- c(3.0, 10.0, 50.0, 100.0, 300.0)
for(tract in Tract.list){
  target_summary <- get(paste("PSJS_HKY_Tract_", toString(tract), "_combined_summary", sep = ""))
  col.names <- target_summary["tract_length", ] < 10*tract
  sim_info <- get(paste("sim.tract.", toString(tract), sep = ""))
  
  hist(target_summary["tract_length", col.names], breaks = 50,
       main = paste("PSJS Estimated Tract length 1/p, Tract = ", toString(tract), ".0 ", sep = ""))
  abline(v =  realized.tract.dist(492, 1.0/tract)$mean, col = "blue")
  abline(v =  tract, col = 2)
  abline(v =  mean(sim_info["mean subtract length", ]), col = "green")
  
  hist(-log(target_summary["tract_length", col.names]), breaks = 50,
       main = paste("PSJS Estimated ln(p), Tract = ", toString(tract), ".0", sep = ""))
  abline(v = log(1.0 / realized.tract.dist(492, 1.0/tract)$mean), col = "blue")
  abline(v = log(1.0 / tract), col = 2)
  abline(v = log(1.0 / mean(sim_info["mean subtract length", ])), col = "green")
  
  
  hist((1/target_summary["tract_length", col.names]), breaks = 50,
       main = paste("PSJS Estimated p, Tract = ", toString(tract), ".0", sep = ""))
  abline(v = 1.0 / tract, col = 2)
  abline(v = 1.0 / realized.tract.dist(492, 1.0/tract)$mean, col = "blue")
  abline(v = 1.0 / mean(sim_info["mean subtract length", ]), col = "green")
  
  plot(sim_info["num IGC", ], sim_info["mean potential tract length", ], 
       main = paste("Simulated potential tract length, Tract = ", toString(tract), sep = ""), 
       xlab = "number of IGC events", ylab = "mean potential tract length")
  abline(h = tract, col = "red")
  abline(h = mean(sim_info["mean subtract length", ]), col = "green")
  
  plot(sim_info["num IGC", ], sim_info["mean realized tract length", ], 
       main = paste("Simulated realized tract length, Tract = ", toString(tract), sep = ""),
       xlab = "number of IGC events", ylab = "mean realized tract length")
  abline(h = realized.tract.dist(492, 1/tract)$mean, col = "red")
  abline(h = mean(sim_info["mean subtract length", ]), col = "green")
  
  plot(sim_info["num IGC", colnames(target_summary)[col.names]], target_summary["tract_length", col.names],
       main = paste("PSJS estimate of Tract ", toString(tract), sep = ""),
       xlab = "number of IGC events", ylab = "PSJS estimated tract length")
  abline(h = tract, col = "red")
  abline(h = realized.tract.dist(492, 1/tract)$mean, col = "blue")
  abline(h = mean(sim_info["mean subtract length", ]), col = "green")
  
  plot(sim_info["num IGC", colnames(target_summary)[col.names]], target_summary["tract_length", col.names],
       main = paste("PSJS estimate of Tract ", toString(tract), sep = ""),
       xlab = "number of IGC events", ylab = "PSJS estimated tract length")
  abline(h = tract, col = "red")
  abline(h = realized.tract.dist(492, 1/tract)$mean, col = "blue")
  abline(h = mean(sim_info["mean subtract length", ]), col = "green")

  plot(sim_info["num IGC with at least two variant sites", colnames(target_summary)[col.names]], 
       target_summary["tract_length", col.names],
       main = paste(" PSJS estimated vs num IGC >1 variant sites - Tract ", toString(tract), sep = ""),
       xlab = "num IGC with at least two variant sites", ylab = "PSJS estimated tract length")
  abline(h = tract, col = 2)
  abline(h = mean(sim_info["mean subtract length", ]), col = "green")
  
  # Now plot kappa estimates
  plot(sim_info["num IGC", colnames(target_summary)[col.names]], 
       target_summary["kappa", col.names],
       main = paste("PSJS kappa, Tract = ", toString(tract), sep = ""),
       xlab = "Num IGC", ylab = "PSJS estimated kappa")
  abline(h = 14.46399, col = 2)
  
  # Now plot r2 estimates
  plot(sim_info["num IGC", colnames(target_summary)[col.names]], 
       target_summary["r2", col.names],
       main = paste("PSJS r2, Tract = ", toString(tract), sep = ""),
       xlab = "Num IGC", ylab = "PSJS estimated r2")
  abline(h = 0.5391702, col = 2)
  
  # Now plot r3 estimates
  plot(sim_info["num IGC", colnames(target_summary)[col.names]], 
       target_summary["r3", col.names],
       main = paste("PSJS r3, Tract = ", toString(tract), sep = ""),
       xlab = "Num IGC", ylab = "PSJS estimated r3")
  abline(h = 11.58006, col = 2)
  
  # Now plot initiation rate estimates
  plot(sim_info["num IGC", colnames(target_summary)[col.names]], 
       target_summary["init_rate", col.names],
       main = paste("PSJS init_rate, Tract = ", toString(tract), sep = ""),
       xlab = "Num IGC", ylab = "PSJS estimated init_rate")
  abline(h = 22.58153 / tract, col = 2)
  
  # Now plot tract p estimates
  plot(sim_info["num IGC", colnames(target_summary)[col.names]], 
       1.0/target_summary["tract_length", col.names],
       main = paste("PSJS tract_p, Tract = ", toString(tract), sep = ""),
       xlab = "Num IGC", ylab = "PSJS estimated tract_p")
  abline(h = 1.0 / tract, col = 2)
  
  # Now plot Tau estimates
  plot(sim_info["num IGC", colnames(target_summary)[col.names]], 
       target_summary["init_rate", col.names]*target_summary["tract_length", col.names],
       main = paste("PSJS Tau, Tract = ", toString(tract), sep = ""),
       xlab = "Num IGC", ylab = "PSJS estimated Tau")
  abline(h = 22.58153, col = 2) 
  
  plot(sim_info["num IGC", colnames(target_summary)[col.names]], 
       target_summary["init_rate", col.names]*target_summary["tract_length", col.names]*3.0/(1.0 +target_summary["r2", col.names] + target_summary["r3", col.names]),
       main = paste("PSJS Weighted Tau, Tract = ", toString(tract), sep = ""),
       xlab = "Num IGC", ylab = "PSJS Tau*3/(1+r2+r3)")
  abline(h = 5.16376333125, col = 2) 
  
  # Now plot initiation rate vs tract length
  plot(target_summary["tract_length", col.names], 
       target_summary["init_rate", col.names],
       main = paste("PSJS init_rate, Tract = ", toString(tract), sep = ""),
       xlab = "Tract length", ylab = "initiation rate")
  lines(sort(target_summary["tract_length", col.names]), 22.58153/sort(target_summary["tract_length", col.names]), 
        col = "red", type = "l")  
  
  plot(sim_info["mean realized tract length", colnames(target_summary)[col.names]], 
       target_summary["tract_length", col.names],
       main = paste(" mean realized vs PSJS estimated Tract ", toString(tract), sep = ""),
       xlab = "mean realized IGC tract length", ylab = "PSJS estimated")
  abline(a= 0.0, b = 1.0, col = "red") 
  
  plot(sim_info["mean potential tract length", colnames(target_summary)[col.names]], 
       target_summary["tract_length", col.names],
       main = paste(" mean potential vs PSJS estimated Tract ", toString(tract), sep = ""),
       xlab = "mean potential IGC tract length", ylab = "PSJS estimated")
  abline(a = 0.0, b = 1.0, col = 2) 
  
  plot(sim_info["mean subtract length", colnames(target_summary)[col.names]], 
       target_summary["tract_length", col.names],
       main = paste(" mean longest variant pair length vs PSJS estimated Tract ", toString(tract), sep = ""),
       xlab = "mean longest variant pair length", ylab = "PSJS estimated")
  abline(a = 0.0, b = 1.0, col = "green") 
  
  
  print(paste("Tract = ", toString(tract), ".0 combined PSJS HKY Results", sep = ""))
  print(matrix(c("Total samples", sum(col.names),
                 "mean estimates", mean(target_summary["tract_length", col.names]), 
                 "sd estimates", sd(target_summary["tract_length", col.names]),
                 "mean longest variant pair", mean(sim_info["mean subtract length", col.names]),
                 "sd longest variant pair", sd(sim_info["mean subtract length", col.names])), 2, 5))  
  
}

```

Now look back at the simulation study of published work to see if Tau estimate is linear with # IGC

```{r}
IGC.path <- "/Users/xji3/FromCluster_IGCCodonSimulation01262016/"
paml.path <- "/Users/xji3/GitFolders/IGCCodonSimulation/"
IGC.geo.list <- c(3.0, 10.0, 50.0, 100.0, 500.0)
num.sim <- 100
paralog <- "YDR418W_YEL054C"
sim.path <- paste("SimulationSummary", paralog, "",sep = "/")
for (IGC.geo in IGC.geo.list){
  summary_mat <- NULL
  IGC.geo.path <- paste("IGCgeo_", toString(IGC.geo), ".0/", sep = "")
  file.prefix <- paste(paralog, "MG94_geo", paste(toString(IGC.geo), ".0", sep = ""), "Sim", sep = "_")
  file.suffix <- "summary.txt"
  for (sim.num in 0:(num.sim - 1)){
    file.name <- paste(file.prefix, toString(sim.num), file.suffix, sep = "_")
    summary_file <- paste(IGC.path, sim.path, IGC.geo.path, file.name, sep = "")
    if (file.exists(summary_file)){
      all <- readLines(summary_file, n = -1)
      col.names <- paste("geo", paste(toString(IGC.geo), ".0", sep = ""), "sim", toString(sim.num), sep = "_")
      row.names <- strsplit(all[length(all)], ' ')[[1]][-1]
      summary_mat <- cbind(summary_mat, as.matrix(read.table(summary_file, 
                                                             row.names = row.names, 
                                                             col.names = col.names)))        
    }
  }
  assign(paste("geo", paste(toString(IGC.geo), ".0", sep = ""), "summary", sep = "_"), summary_mat)
}

# Now read in Total number of IGC events in simulation
edge.list <- c(#"N0_kluyveri", outgroup branch has no IGC events
  "N0_N1", "N1_castellii", "N1_N2", "N2_bayanus", 
  "N2_N3", "N3_kudriavzevii", "N3_N4", "N4_mikatae", "N4_N5", 
  "N5_cerevisiae", "N5_paradoxus")
nIGC.mat <- NULL
for(IGC.geo in IGC.geo.list){
  summary_mat <- NULL
  IGC.geo.path <- paste("IGCgeo_", toString(IGC.geo), ".0/sim_", sep = "")
  for(sim.num in 0:(num.sim - 1)){
    num.IGC <- 0
    for (edge in edge.list){
      file.name <- paste(IGC.geo.path, toString(sim.num), "/log/", edge, "_log.log", sep = "")
      log.file <- paste(IGC.path, paralog, "/", file.name, sep = "")
      content <- readLines(log.file)
      if (length(content)>6){
        for(i in 7:length(content)){
          line.cont <- strsplit(content[i], " ")
          if(line.cont[[1]][1] == "IGC"){
            num.IGC <- num.IGC + 1
          }
        }
      }
      
    }
    summary_mat <- c(summary_mat, num.IGC)
  }
  nIGC.mat <- rbind(nIGC.mat, summary_mat)
}
rownames(nIGC.mat) <- paste("IGCgeo_", IGC.geo.list, ".0", sep = "")
colnames(nIGC.mat) <- paste("sim_", 0:99, sep = "")

for(IGC.geo in IGC.geo.list){
  summary_mat <- get(paste("geo_", IGC.geo, ".0_summary", sep = ""))
  # Now plot Tau estimates
  plot(nIGC.mat[paste("IGCgeo_", IGC.geo, ".0", sep = ""),], 
       summary_mat["tau", ],
       main = paste("PSJS Tau, Tract = ", toString(IGC.geo), sep = ""),
       xlab = "Num IGC", ylab = "IS MG94 estimated Tau")
  abline(h = 1.409408, col = 2)   
  
}

```


```{r}
for(tract in Tract.list){
  sim.info <- get(paste("sim.tract.", toString(tract), sep = ""))
  PSJS.info <- get(paste("PSJS_HKY_Tract_", toString(tract), "_combined_summary", sep = ""))
  shared.col <- colnames(PSJS.info)[PSJS.info["tract_length", ] < 10000 & PSJS.info["tract_length", ] > 1.]
  
  # Now show the ratio of PSJS estimated tract / actual mean tracts in simulation
  PSJS.ratio <- PSJS.info["tract_length", shared.col]/sim.info[1, shared.col]
  hist(PSJS.ratio, breaks = 50, main = paste("PSJS ratio Tract = ", toString(tract), sep = ""))
  print(matrix(c("Tract", tract, "PSJS mean", mean(PSJS.ratio), "PSJS sd", sd(PSJS.ratio)), 2, 3))
}

```

save workspace now

```{r}
save.image("./SimulationStudy.RData")
```