---
title: "QualityControl"
author: "Xiang Ji"
date: "11/16/2017"
output: html_document
---

This file summarizes the data sets in Harpak et al. 2017. Main interests are 
length of alignment, level of missing species etc.

```{r}
rm(list=ls())  # clean up workspace
setwd("/Users/xji3/GitFolders/YeastIGCTract/Harpak/")
alignment.files <- list.files("./intronAlignmentSeparated/")
filtered.alignment.files <- NULL
num.missing.genes.list <- NULL
filtered.files <- NULL
for(seq.file in alignment.files){
  info <- read.table(paste("./intronAlignmentSeparated/", seq.file, sep = ""), as.is = TRUE)
  num.missing.genes <- sum(is.na(info[, 1]))
  num.missing.genes.list <- c(num.missing.genes.list, num.missing.genes)
  if(num.missing.genes == 0){
    filtered.files <- c(filtered.files, seq.file)
  }
  if(dim(info)[2] > 1){
    filtered.alignment.files <- c(filtered.alignment.files, seq.file)
  }
}
table(num.missing.genes.list)
```

OK, 27 datasets have no missing genes (based on gene names). Now apply additional filter by looking at the gap percentage (100% gaps = missing).

```{r}
additional.failure <- NULL
for(seq.file in filtered.files){
  info <- read.table(paste("./intronAlignmentSeparated/", seq.file, sep = ""), as.is = TRUE)
  for(row.num in 1:11){
    seq <- strsplit(info[row.num, 2], "")[[1]]
    percent.gap <- (length(grep("-", seq)) + length(grep("N", seq)))/length(seq)
    if (percent.gap > 1 - 1e-6){
      if( ! seq.file %in% additional.failure)
        additional.failure <- c(additional.failure, seq.file)
    }
  }
}
additional.filtered <- filtered.files[! filtered.files %in% additional.failure]
# See how many have missing info
length(additional.failure)
# See how many have passed the additional filter
length(additional.filtered)
# Show the remaining data sets
additional.filtered
```

Now summarize all alignment files in terms of how many ingroup species have missing genes.

```{r}
num.missing.species.list  <- NULL
for(num.spe in 0:5){
  assign(paste("missing", num.spe, "species", sep = "."), NULL)
}
for(seq.file in filtered.alignment.files){
  info <- read.table(paste("./intronAlignmentSeparated/", seq.file, sep = ""), as.is = TRUE)
  for(row.num in 1:11){
    seq <- strsplit(info[row.num, 2], "")[[1]]
    percent.gap <- length(grep("-", seq))/length(seq)
    if (percent.gap > 1 - 1e-6){
      info[row.num, 1] <- NA
    }   
  }
  num.spe <- 0
  for(row.iter in 1:5){
    if(any(is.na(info[c(row.iter*2-1, row.iter*2), 1]))){
      num.spe <- num.spe + 1
    }
  }
  num.missing.species.list <- c(num.missing.species.list, num.spe)
  assign(paste("missing", num.spe, "species", sep = "."), c(get(paste("missing", num.spe, "species", sep = ".")), seq.file))
}
table(num.missing.species.list)
# In addition, show files that have only one ingroup species that misses either one or both copies
missing.1.species
```

Now look at the length of the alignment and percent of columns that contain gaps

```{r}
# 20 files that have no missing genes
missing.0.summary <- NULL
for(seq.file in missing.0.species){
  info <- read.table(paste("./intronAlignmentSeparated/", seq.file, sep = ""), as.is = TRUE)

  gap.col <- NULL
  for(row.num in 1:11){
    seq <- strsplit(info[row.num, 2], "")[[1]]
    gap.col <- union(gap.col, grep("-", seq))
  }
  
  missing.0.summary <- rbind(missing.0.summary, c(length(seq), length(seq) - length(gap.col), length(gap.col), length(gap.col)/length(seq)))
}
rownames(missing.0.summary) <- missing.0.species
colnames(missing.0.summary) <- c("length", "# col no gap", "# col with gap", "% gap col")
missing.0.summary

# 46 files that have one species that contains missing genes
missing.1.summary <- NULL
for(seq.file in missing.1.species){
  info <- read.table(paste("./intronAlignmentSeparated/", seq.file, sep = ""), as.is = TRUE)

  gap.col <- NULL
  for(row.num in 1:11){
    seq <- strsplit(info[row.num, 2], "")[[1]]
    if(length(grep("-", seq)) < length(seq)){
      gap.col <- union(gap.col, grep("-", seq))
    }
  }
  
  missing.1.summary <- rbind(missing.1.summary, c(length(seq), length(seq) - length(gap.col), length(gap.col), length(gap.col)/length(seq)))
}
rownames(missing.1.summary) <- missing.1.species
colnames(missing.1.summary) <- c("length", "# col no gap", "# col with gap", "% gap col")
missing.1.summary

# Now save the file names
write.table(missing.0.species, "./missing_0_species_list.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)
write.table(missing.1.species, "./missing_1_species_list.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)
```